Effectiveness of community-based TURF-reserves in Mexican small-scale fisheries

### 2019-04-07 docs/PLoS_manuscript.tex
PLoS Submission ready

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Effectiveness of community-based TURF-reserves in Mexican small-scale fisheries} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Juan Carlos Villaseñor-Derbez\textsuperscript{1\Yinyang*},
Eréndira Aceves-Bueno\textsuperscript{1,2\Yinyang},
Stuart Fulton\textsuperscript{3\ddag},
Álvin Suarez\textsuperscript{3\ddag},
Arturo Hernández-Velasco\textsuperscript{3\ddag},
Jorge Torre\textsuperscript{3\ddag},
Fiorenza Micheli\textsuperscript{4\ddag}
\\
\bigskip
\textbf{1} Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA
\\
\textbf{2} Nicholas School of the Environment, Duke University, Beaufort, NC, USA
\\
\textbf{3} Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico
\\
\textbf{4} Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA
\\
\bigskip

@@

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world. Small-scale fisheries face great challenges since they are difficult to monitor, enforce, and manage. Combining territorial use rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments. In the last 13 years, 18 old and new community-based Mexican TURF-reserves gained legal recognition thanks to a regulation passed in 2012; their effectiveness has not been formally evaluated. We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico. We find that reserves have not yet achieved their stated goal of increasing the density of lobster and other benthic invertebrates. A lack of clear ecological and socioeconomic effects likely results from a combination of factors. First, some of these reserves might be too young for the effects to show. Second, the reserves are not large enough to protect mobile species, like lobster. Third, variable and extreme oceanographic conditions have impacted harvested populations. Fourth, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect in catches. However, these reserves may provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species.

@@

Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \cite{pauly_2005-qV,worm_2006-IB,halpern_2008-dK}. In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce \cite{costello_2012}. One of the many approaches taken to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves, thus creating TURF-reserve systems \cite{afflerbach_2014,gelcich_2015,lester_2017}.

@@

Here, we combine causal inference techniques and the Social-Ecological Systems (SES) framework to evaluate community-based TURF-reserves in three coastal communities in Mexico. The objective of this work is twofold. First, to provide a holistic evaluation of the effectiveness of community-based TURF-reserves in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop, which may inform similar processes in other countries. Second, to identify opportunities where improvement or adjustment might lead to increased effectiveness of these reserves.
%We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based TURF-reserves in small-scale fisheries where this tool is used to manage and rebuild coastal fisheries.

@@

The most common setup of community-based TURF-reserves in Mexico is the following. Fishers from a given community are assembled in fishing cooperatives which have exclusive fishing rights over a spatially delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig~\ref{fig:map}A). Each TURF is exclusively fished by one cooperative, and each community usually hosts no more than one cooperative. The profits from each TURF are shared amongst all fishers from the cooperative. Fishing cooperatives interested in implementing marine reserves within their TURFs (\emph{i.e.} TURF-reserves) work with CSOs to design them. Fishers then ask the government to grant legal recognition to their TURF-reserves as Fish Refuges under the 2012 regulation \cite{nom}.

@@

We evaluate three community-based no-take TURF-reserve systems implemented in Mexican TURF-managed fisheries (Fig~\ref{fig:map}A). The first one was created by the \emph{Buzos y Pescadores de la Baja California} fishing cooperative, located in Isla Natividad in the Baja California Peninsula (Fig~\ref{fig:map}B). The main fishery in the island is the spiny lobster (\emph{Panulirus interruptus}), but other resources like finfish, sea cucumber, sea urchin, snail, and abalone are also an important source of income. In 2006, the community decided to implement two marine reserves within their fishing grounds. The objective of these reserves was "to protect and recover stocks of commercially important invertebrate species"; mainly lobster and abalone. The reserves obtained legal recognition in 2018 \cite{dof_website_2018}.

The other two TURF-reserve systems are located in Maria Elena and Punta Herrero, in the Yucatan Peninsula (Fig~\ref{fig:map}C). In contrast with Isla Nativdad, which hosts a well-established fishing community, Maria Elena is a fishing camp visited intermittently during the fishing season that belongs to the \emph{Cozumel} fishing cooperative. Punta Herrero is home to the \emph{José María Azcorra} fishing cooperative, and similar to Isla Natividad hosts a small community. Their main fishery is the Caribbean spiny lobster (\emph{Panulirus argus}), but they also target finfish in the off-season. Maria Elena and Punta Herrero established eight and four marine reserves in 2012 and 2013, respectively. These reserves have been legally recognized as Fishing Refuges since their original implementation \cite{dof_website_2012,dof_website_2013} and subsequent re-establishments \cite{dof_website_2017b}.

@@

The regulation governing the implementation of Fish Refuges states that these are fishery management tools intended to have conservation and fisheries benefits \cite{nom}. For this reason, the main portion of our analyses focuses on a series of biological and socioeconomic indicators that may respond to reserve implementation. However, the effectiveness of conservation and fisheries management interventions also depends on the social and governance structures in place. We therefore incorporate a reduced version of the Social Ecological Systems framework \cite{ostrom_2009} and evaluate variables and indicators known to aid and hinder the effectiveness of management interventions in conservation and fisheries. The incorporation of the SES is not intended to relate different levels of governance with reserve effectiveness, but rather help provide context on the social-ecological system in which reserves develop. The following two sections describe our data collection methods and analyses.

@@

We use three main sources of information to evaluate these reserves across ecological, socioeconomic, and governance dimensions. Ecological data come from the annual ecological monitoring of reserve and control sites. Reserve sites are areas where no fishing occurs. Control sites are areas that meet the following criteria: i) habitat characteristics are similar to the corresponding reserves, ii) presumably had a similar probability of being selected as reserves during the design phase, iii) are located within the TURF, where fishing occurs, and iv) are not directly adjacent to the reserves. We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve. This provides us with a Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \cite{stewartoaten_1986,depalma_2018,kerr_2019} and causally attribute the changes to the reserve \cite{francinifilho_2008,Villasenor-Derbez_2018}.

The biological data are collected by members from each community and personnel from the Mexican CSO \emph{Comunidad y Biodiversidad} (\href{https://www.cobi.org.mx}{COBI}). Trained divers record species richness and abundances of fish and invertebrate species along replicate transects (30 $\times$ 2 m each) at depths 5-20 m in the reserves and control sites \cite{suman_2010-ez,fulton_2018,fulton_2019}. Size structures are also collected during fish surveys. All sites were surveyed annually, and at least once before implementation of the reserves. A summary of sampling effort and time series are shown in the supplementary materials (\nameref{S1_Table}, \nameref{S2_Table}, \nameref{S1_Fig}, \nameref{S2_Fig}, \nameref{S3_Fig}, \nameref{S4_Fig}).

Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings (Kg) and revenues (MXP) for TURF-managed cooperatives with and without marine reserves. In this case our treated unit are the cooperatives (\emph{i.e.} communities) that have implemented a reserve within their TURF, and the controls are adjacent communities that have a TURF but did not implement a reserve. Cooperatives incorporated in this analysis have similar number of members, belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \cite{mccay_2014,mccay_2017,ayer_2018}. Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \cite{oecd_2017}. A table with summary statistics for this data is provided in the supplementary materials (\nameref{S3_Table},\nameref{S5_Fig}).

@@

\input{tab/ses.tex}

@@

\input{tab/indicators.tex}

@@

Where year-level fixed effects capturing a temporal trend are represented by $\gamma_t Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites. The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the treated sites at time $t$ after controlling for other time and space variations (\emph{i.e.} $\gamma_t$ and $\beta$ respectively). Therefore, we would expect this term to be positive if the indicator increases because of the reserve. Finally, $\epsilon_{i,t}$ represents the error term of the regression.

@@

The coefficient interpretations remains as for Eq. \ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$). These regression models allow us to establish a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and site-specific dynamics \cite{depalma_2018}. Since we are interested in the effectiveness of each reserve system, we fit one model for each indicator in each community (\emph{e.g.} there are three models for lobster density, one for each community). This gives us a total of 12 biological model fits and four socioeconomic model fits. Model coefficients were estimated via ordinary least-squares and used heteroskedastic-robust standard error correction \cite{zeileis_2004-7n}. All analyses were performed in R 3.5.2 and R Studio version 1.1.456 \cite{R_2018}. All data and code needed to reproduce our analyses are available in a GitHub repository at: \url{https://github.com/jcvdav/ReserveEffect}.

TURF-reserve systems are inherently intricate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \cite{ostrom_2009,gelcich_2015}. When evaluating the effects of TURF-reserves, it is important to consider not only the indicators of interest, but also the governance settings under which a reserve operates. We use the SES framework to qualitatively evaluate each community and create a narrative that provides context for each of them. The use of this framework standardizes our analysis and allows us to communicate our results in a common language across fields by using a set of previously defined variables and indicators. Due to the lack of sufficient information to quantitatively operationalize the social-ecological systems framework for these case studies (as done in \cite{leslie_2015-na}), we followed a similar approach to \cite{basurto_2013-oq}, who used the SES framework as a classification system of the available information to qualitatively analyze fisheries systems. We based our variable selection primarily on previous analyses of Mexican fishing cooperatives \cite{leslie_2015-na,basurto_2013-oq}. We also incorporate other relevant variables known to influence reserve performance following, such as age and size of reserve or isolation of the system \cite{difranco_2016-Xw,edgar_2014-UO}. Table~\ref{table:ses} shows the selected variables, along with definitions and values.

@@

Indicators showed ambiguous responses through time for each reserve. Fig~\ref{fig:indicators}A shows positive effect sizes for lobster densities in Isla Natividad and Punta Herrero during the first years, but the effect is eroded through time. In the case of Maria Elena, positive changes were observed in the second and third years. These effects were in the order of 0.01 extra organisms $\mathrm{m}^{-2}$, but were only significantly different from zero for Maria Elena ($p < 0.05$). Likewise, no significant changes were detected in fish biomass or invertebrate and fish densities (Fig. \ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends. Figures and tables with time series of indicators and model coefficients are presented in the supplementary materials (\nameref{S1_Fig}, \nameref{S2_Fig}, \nameref{S3_Fig}, \nameref{S4_Fig}, \nameref{S4_Table}, \nameref{S5_Table}, \nameref{S6_Table}).

\begin{figure}[!h]
\centering
%\includegraphics{img/fig_bio_results.pdf}
\caption{{\bf Effect sizes for biological indicators}
Points indicate the effect size and error bars are heteroskedastic-robust standard errors. Years have been centered to year of implementation. Colors and shapes denote communities: Isla Natividad (IN; red circles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares). Letters denote indicator: A) lobster densities (\emph{Panulirus spp}, B) fish biomass, C) invertebrate densities, and D) fish densities. Plots are ordered by survey type (left column: invertebrates; right column: fish). Points are jittered hotizontally to avoid overplotting.}
\label{fig:indicators}
\end{figure}

@@

Lobster landings and revenue were only available for Isla Natividad and Maria Elena (Fig~\ref{fig:lobsters}). For all years before implementation, the effect sizes are close to zero, indicating that the control and treatment sites have similar pre-treatment trends, suggesting that these are plausible controls. However, effect sizes do not change after the implementation of the reserve. Interestingly, the negative effect observed for Isla Natividad on year 5 corresponds to the 2011 hypoxia events \cite{micheli_2012-EU}. The only positive change observed in lobster landings is for Isla Natividad in 2014 ($p < 0.1$). The year of post-implementation data for Maria Elena does not show a significant effect of the reserve. Isla Natividad shows higher revenues after the implementation of the reserve, as compared to the control communities. However, these changes are only significant for the third year ($p < 0.05$). Full tables with model coefficients are presented in the supplementary materials (\nameref{S7_Table}, \nameref{S8_Table}).

@@

Our results indicate that these TURF-reserves have not increased lobster densities. Additionally, no co-benefits were identified when using other ecological indicators aside from the previously reported buffering effect that reserves can have to environmental variability in Isla Natividad \cite{micheli_2012-EU}. The socioeconomic indicators pertaining landings and revenues showed little to no change after reserve implementation. Lastly, the communities exhibit all the social enabling conditions for effective reserve and resource management. Here we discuss possible shortcomings in our analyses as well alternative explanations for the observed lack of effectiveness.

@@

While the use of causal inference techniques may help us support evidence-based conservation, spatial connectivity between reserve and control sites, stockpiling, and backstopping may confound the results \cite{kerr_2019}. Given that we find no clear evidence of reserve effectiveness, one might say that our reserve and control sites are not spatially independent. This would imply that the recovery within the reserve quickly results in recovery ouside the protected area. However, indicators show little to no temporal variation (\nameref{S1_Fig}, \nameref{S2_Fig}, \nameref{S3_Fig}, \nameref{S4_Fig}), and it is unlikely that this effect would be observed under current reserve designs, as detailed below.

>>>>

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Effectiveness of community-based TURF-reserves in Mexican small-scale fisheries} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Juan Carlos Villaseñor-Derbez\textsuperscript{1\Yinyang*},
Eréndira Aceves-Bueno\textsuperscript{1,2\Yinyang},
Stuart Fulton\textsuperscript{3\ddag},
Alvin Suarez\textsuperscript{3\ddag},
Arturo Hernández-Velasco\textsuperscript{3\ddag},
Jorge Torre\textsuperscript{3\ddag},
Fiorenza Micheli\textsuperscript{4\ddag}
\\
\bigskip
\textbf{1} Bren School of Environmental Science \& Management, University of California, Santa Barbara, Santa Barbara, CA, USA
\\
\textbf{2} Nicholas School of the Environment, Duke University, Beaufort, NC, USA
\\
\textbf{3} Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico
\\
\textbf{4} Hopkins Marine Station and Stanford Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA
\\
\bigskip

@@

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world. Small-scale fisheries face great challenges since they are difficult to monitor, enforce, and manage. Combining territorial use rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments. Since 2012, 18 old and new community-based Mexican TURF-reserves gained legal recognition thanks to a regulation passed in 2012; their effectiveness has not been formally evaluated. We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico. We find that, overall, reserves have not yet achieved their stated goals of increasing the density of lobster and other benthic invertebrates, nor increasing lobster catches. A lack of clear ecological and socioeconomic effects likely results from a combination of factors. First, some of these reserves might be too young for the effects to show. Second, the reserves are not large enough to protect mobile species, like lobster. Third, variable and extreme oceanographic conditions have impacted harvested populations. Fourth, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect in catches. However, even small reserves are expected to provide benefits for sedentary invertebrates over longer time frames, with continued protection. These reserves may provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species.

@@

Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \cite{pauly_2005-qV,worm_2006-IB,halpern_2008-dK}. In particular, small-scale fisheries face great challenges since they tend to be hard to monitor and enforce \cite{costello_2012}. One of the many approaches taken to improve the performance of coastal fisheries and health of the local resources is through the implementation of Territorial Use Rights for Fisheries (TURFs) that contain no-take marine reserves\cite{afflerbach_2014,gelcich_2015,lester_2017}.

@@

Here, we combine causal inference techniques and the Social-Ecological Systems (SES) framework to evaluate community-based TURF-reserves in three coastal communities in Mexico. The objective of this work is twofold. First, to provide a holistic evaluation of the effectiveness of community-based TURF-reserves in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop, which may inform similar processes in other countries. The second objective is to identify opportunities where improvement or adjustment might lead to increased effectiveness of these reserves.
%We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based TURF-reserves in small-scale fisheries where this tool is used to manage and rebuild coastal fisheries.

@@

The most common setup of community-based TURF-reserves in Mexico is the following. Fishers from a given community assemble into a fishing cooperative which has exclusive fishing rights over a spatially delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig~\ref{fig:map}A). Each TURF is exclusively fished by one cooperative, and each community usually hosts no more than one cooperative. The profits from each TURF are shared amongst all fishers from the cooperative. Fishing cooperatives interested in implementing marine reserves within their TURFs (\emph{i.e.} TURF-reserves) work with CSOs to design them. Fishers then ask the government to grant legal recognition to their TURF-reserves as Fish Refuges, as stated in the 2014 regulation \cite{nom}.

@@

We evaluate three community-based no-take TURF-reserve systems implemented in Mexican TURF-managed fisheries (Fig~\ref{fig:map}A). The first one was created by the \emph{Buzos y Pescadores de la Baja California} fishing cooperative, located in Isla Natividad in the Baja California Peninsula (Fig~\ref{fig:map}B). At present, the main fishery in the island is the spiny lobster (\emph{Panulirus interruptus}), but other resources like finfish, sea cucumber, sea urchin, snail, and abalone are also an important source of income. In 2006, the community decided to implement two marine reserves within their fishing grounds. The objective of these reserves was "to protect and recover stocks of commercially important invertebrate species"; mainly lobster and abalone. The reserves obtained legal recognition in 2018 \cite{dof_website_2018}.

The other two TURF-reserve systems are located in Maria Elena and Punta Herrero, in the Yucatan Peninsula (Fig~\ref{fig:map}C). In contrast with Isla Nativdad, which hosts a well-established fishing community, Maria Elena is a fishing camp visited intermittently during the fishing season that belongs to the \emph{Cozumel} fishing cooperative. Punta Herrero is home to the \emph{José María Azcorra} fishing cooperative, and similar to Isla Natividad hosts a small community. Their main fishery is the Caribbean spiny lobster (\emph{Panulirus argus}), but they also target finfish in the off-season. Maria Elena and Punta Herrero established eight and four marine reserves in 2012 and 2013, respectively. These reserves have been legally recognized as Fishing Refuges since their original implementation \cite{dof_website_2012,dof_website_2013} and subsequent re-establishments in 2017 \cite{dof_website_2017b}.

@@

The regulation governing the implementation of Fish Refuges states that these are fishery management tools intended to have conservation and fisheries benefits \cite{nom}. For this reason, the main portion of our analyses focuses on a series of biological and socioeconomic indicators that may respond to reserve implementation. However, the effectiveness of conservation and fisheries management interventions also depends on the social and governance structures in place. We therefore incorporate a reduced version of the Social Ecological Systems framework \cite{ostrom_2009} and evaluate variables and indicators known to aid and hinder the effectiveness of management interventions in conservation and fisheries. The incorporation of the SES is not intended to relate different levels of governance with reserve effectiveness, but rather to provide context on the social-ecological system in which reserves develop. The following two sections describe our data collection methods and analyses.

@@

We use three main sources of information to evaluate these reserves across ecological, socioeconomic, and governance dimensions. Ecological data come from the annual ecological monitoring of reserve and control sites. Reserve sites are areas where no fishing occurs. Control sites are areas that meet the following criteria: i) habitat characteristics are similar to the corresponding reserves, ii) presumably had a similar probability of being selected as reserves during the design phase, iii) are located within the TURF, where fishing occurs, and iv) are not directly adjacent to the reserves. We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve. This provides us with a Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that allows us to capture and control for temporal and spatial dynamics and causally attribute the changes to the reserve \cite{stewartoaten_1986,francinifilho_2008,depalma_2018,kerr_2019,Villasenor-Derbez_2018}.

The biological data are collected by members from each community and personnel from the Mexican CSO \emph{Comunidad y Biodiversidad} (\href{https://www.cobi.org.mx}{COBI}). Trained divers record species richness and abundances of fish and invertebrate species along replicate transects (30 $\times$ 2 m each) at depths 5-20 m in the reserves and control sites \cite{suman_2010-ez,fulton_2018,fulton_2019}. Size structures are also collected during fish surveys. All sites were surveyed annually, and at least once before implementation of the reserves. A summary of sampling effort and time series are shown in the supplementary materials (Tables A-B and Figures A-D in \nameref{S1_Text}).

Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings (Kg) and revenues (MXP) for TURF-managed cooperatives with and without marine reserves. In this case our treated unit are the cooperatives (\emph{i.e.} communities) that have implemented a reserve within their TURF, and the controls are adjacent communities that have a TURF but did not implement a reserve. Cooperatives incorporated in this analysis have similar number of members, belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \cite{mccay_2014,mccay_2017,ayer_2018}. Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \cite{oecd_2017}. A table with summary statistics and time series for this data are provided in the supplementary materials (Table C and Figure E in \nameref{S1_Text}).

@@

\begin{table}[!ht]

\caption{{\bf Variables for the Social-Ecological System analysis.}}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{>{\raggedright\arraybackslash}p{6.5cm}|>{\raggedright\arraybackslash}p{12.2cm}}
\hline
Variable & Narrative\\
\hline
\multicolumn{2}{l}{\textbf{Resource System (RS)}}\\
\hline
\hspace{1em}RS2 - Clarity of system boundaries: Clarity of geographical boundaries of TURF and reserves & Individual TURF and reserve boundaries are explicitly outlined in official documents that include maps and coordinates. Reserve placement is decided by the community. Fishers use GPS units and landmarks.\\
\hline
\hspace{1em}RS3 - Size of resource system: TURF Area (Km$^2$) & IN = 889.5; ME = 353.1; PH = 299.7\\
\hline
\hspace{1em}RS3 - Size of resource system: Reserve area (Evaluated reserve area; Km$^2$) & IN = 2 (1.3); ME = 10.48(0.09); PH = 11.25 (4.37)\\
\hline
\hspace{1em}RS4.1 - Stock status: Status of the main fishery & Lobster stocks are well managed, and are (IN) or have been (ME, PH) MSC certified.\\
\hline
\hspace{1em}*RS5 - Age of reserves: Years since reserves were implemented & IN = 12; ME = 6; PH = 5\\
\hline
\multicolumn{2}{l}{\textbf{Resource Unit (RU)}}\\
\hline
\hspace{1em}RU1 - Resource unit mobility & Adult spiny lobsters can move between 1 and 10 Km, while larvae can have displacements in the order of hundreds of Km (Aceves-Bueno et al., 2017; Green et al., 2017).\\
\hline
\hspace{1em}RU5 - Number of units (catch diversity): Number of targeted species & Lobster is their main fishery of these three communities, but they also target finfish (2 spp each). Additionally, fishers from Isla Natividad target other sedentary benthic invertebrates (4 spp).\\
\hline
\multicolumn{2}{l}{\textbf{Actors (A)}}\\
\hline
\hspace{1em}A1 - Number of relevant actors: Number of fishers & IN = 98; ME = 80; PH = 21\\
\hline
\hspace{1em}*A3 - Isolation: Level of isolation of the fishing grounds & Their fishing grounds and reserves are highly isolated and away from dense urban centers. IN lies 545 Km south from Tijuana, and ME and PH 230 Km south from Cancun, where the nearest international airports are located.\\
\hline
\multicolumn{2}{l}{\textbf{Governance system (G)}}\\
\hline
\hspace{1em}GS6.1.4.3 - Territorial use communal rights : Presence of institutions that grant exclusive harvesting rights & Each community has exclusive access to harvest benthic resources, including lobster. These take the form of Territorial User Rights for Fisheries granted by the government to fishing cooperatives.\\
\hline
\hspace{1em}GS6.2 - Operational rules: Rules implemented by individuals atuhorized to partake on collective activities & Fishers have rules in addition to what the legislation mandates. These are: larger minimum catch sizes, lower quotas, and assigning fishers to specific fishing grounds within their TURF.\\
\hline
\hspace{1em}GS9.1 - Social monitoring: Monitoring of the activities performed by cooperative members and external fishers & Fishing cooperatives have a group (Consejo de vigilancia) that monitors and enforces formal and internal rules. They ensure fishers of their fishing cooperative adhere to the established rules, and that foreign vessels do not poach their TURF and reserves.\\
\hline
\hspace{1em}GS9.2 - Biophysical monitoring: Monitoring of biological resources, including targeted species & Fishers perform annual standardized underwater surveys in the reserves and fishing grounds. Recently, they have installed oceanographic sensors to monitor oceanographic variables.\\
\hline
GS10.1 - Graduated sanctions & Fishers have penalties for breaking collective-choice rules or fishing inside the reserves. These may range from scoldings and warnings to not being allowed to harvest a particular resource or being expelled from the cooperative.\\
\hline
\end{tabular}}
\label{table:ses}
IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero. Alphanumeric codes for variables follow \cite{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \cite{difranco_2016-Xw} and \cite{edgar_2014-UO}. The presented narrative applies equally for all communities unless otherwise noted.
\end{table}

@@

\begin{table}[!ht]

\caption{
{\bf List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}}
\centering
\begin{tabular}{l|l}
\hline
Indicator & Units\\
\hline
\multicolumn{2}{l}{\textbf{Biological}}\\
\hline
\hspace{1em}Lobster density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Invertebrate density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Fish density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Fish biomass & Kg $\mathrm{m}^{-2}$\\
\hline
\multicolumn{2}{l}{\textbf{Socioeconomic}}\\
\hline
\hspace{1em}Income from target species & M MXP\\
\hline
\hspace{1em}Landings from target species & Metric Tonnes\\
\hline
\end{tabular}
\label{table:indicators}
\end{table}

@@

Where year-level fixed effects capturing a temporal trend are represented by $\gamma_t Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites. The effect of the reserve is captured by the $\lambda_t$ vector of coefficients, and represents the difference observed between the control site before the implementation of the reserve and the treated sites at time $t$ after controlling for other time and space variations (\emph{i.e.} $\gamma_t$ and $\beta$ respectively). Therefore, we would expect this term to be positive if the indicator increases because of the reserve. Finally, $\epsilon_{i,t}$ represents the error term of the regression.

@@

The coefficient interpretations remains as for Eq. \ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$).

These regression models allow us to establish a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and site-specific dynamics \cite{depalma_2018}. Since we are interested in the effectiveness of each reserve system, we fit one model for each indicator in each community (\emph{e.g.} there are three models for lobster density, one for each community). This gives us a total of 12 biological model fits and four socioeconomic model fits. Model coefficients were estimated via ordinary least-squares and used heteroskedastic-robust standard error correction \cite{zeileis_2004-7n}. All analyses were performed in R 3.5.2 and R Studio version 1.1.456 \cite{R_2018}. All data and code needed to reproduce our analyses are available in a GitHub repository at: \url{https://github.com/jcvdav/ReserveEffect}.

TURF-reserve systems are inherently intricate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \cite{ostrom_2009,gelcich_2015}. When evaluating the effects of TURF-reserves, it is important to consider not only the indicators of interest, but also the governance settings under which a reserve operates. We use the SES framework to qualitatively evaluate each community and create a narrative that provides context for each of them. The use of this framework standardizes our analysis and allows us to communicate our results in a common language across fields by using a set of previously defined variables and indicators. Due to the lack of sufficient information to quantitatively operationalize the social-ecological systems framework for these case studies (as done in \cite{leslie_2015-na}), we followed a similar approach to \cite{basurto_2013-oq}, who used the SES framework as a classification system of the available information to qualitatively analyze fisheries systems. We based our variable selection primarily on previous analyses of Mexican fishing cooperatives \cite{leslie_2015-na,basurto_2013-oq}. We also incorporate other relevant variables known to influence reserve performance, such as age and size of reserve or isolation of the system \cite{difranco_2016-Xw,edgar_2014-UO}. Table~\ref{table:ses} shows the selected variables, along with definitions and values.

@@

Indicators showed ambiguous responses through time for each reserve. Fig~\ref{fig:indicators}A shows positive effect sizes for lobster densities in Isla Natividad and Punta Herrero during the first years, but the effect is eroded through time. In the case of Maria Elena, positive changes were observed in the second and third years. These effects were in the order of 0.01 extra organisms $\mathrm{m}^{-2}$, but were only significantly different from zero for Maria Elena ($p < 0.05$). Likewise, no significant changes were detected in fish biomass or invertebrate and fish densities (Fig. \ref{fig:indicators}B-D). Invertebrate and fish densities showed positive trends in Natividad in most years. Reserves in Maria Elena and Punta Herrero showed positive trends for invertebrates in some years. However, these were not statistically significant. Figures with time series of indicators and tables with model coefficients and main effects are presented in the supplementary materials (Figures A-D and Tables D-F in \nameref{S1_Text}).

\begin{figure}[h]
\centering
%\includegraphics{img/fig_bio_results.pdf}
\caption{{\bf Effect sizes for biological indicators}
Points indicate the effect size and error bars are heteroskedastic-robust standard errors. Years have been centered to year of implementation. Colors and shapes denote communities: Isla Natividad (IN; red circles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green squares). Letters denote indicator: A) lobster densities (\emph{Panulirus spp}, B) fish biomass, C) invertebrate densities, and D) fish densities. Plots are ordered by survey type (left column: invertebrates; right column: fish). Points are jittered hotizontally to avoid overplotting.}
\label{fig:indicators}
\end{figure}

@@

Lobster landings and revenue were only available for Isla Natividad and Maria Elena (Fig~\ref{fig:lobsters}). For all years before implementation, the effect sizes are close to zero, indicating that the control and treatment sites have similar pre-treatment trends, suggesting that these are plausible controls. However, effect sizes do not change after the implementation of the reserve. Interestingly, the negative effect observed for Isla Natividad on year 5 corresponds to the 2011 hypoxia events \cite{micheli_2012-EU}. The only positive change observed in lobster landings is for Isla Natividad in 2014 ($p < 0.1$). The year of post-implementation data for Maria Elena does not show a significant effect of the reserve. Isla Natividad shows higher revenues after the implementation of the reserve, as compared to the control communities, though these changes are only significant for the third year ($p < 0.05$). Full tables with model coefficients are presented in the supplementary materials (Tables G-H in \nameref{S1_Text}).

@@

Our results indicate that these TURF-reserves have not increased lobster densities. Additionally, no co-benefits were identified when using other ecological indicators aside from the previously reported buffering effect that reserves can have to environmental variability in Isla Natividad, as well as positive trends in invertebrate and fish densities \cite{micheli_2012-EU}. The socioeconomic indicators pertaining landings and revenues showed little to no change after reserve implementation. Lastly, the communities exhibit all the social enabling conditions for effective reserve and resource management. Here we discuss possible shortcomings in our analyses as well alternative explanations for the observed lack of effectiveness.

@@

While the use of causal inference techniques may help us support evidence-based conservation, spatial connectivity between reserve and control sites, stockpiling, and backstopping may confound the results \cite{kerr_2019}. Given that we find no clear evidence of reserve effectiveness, one might say that our reserve and control sites are not spatially independent. This would imply that the recovery within the reserve quickly results in recovery outside the protected area. However, indicators show little to no temporal variation (Figures A-D in \nameref{S1_Text}), and it is unlikely that this effect would be observed under current reserve designs, as detailed below.
